### 2015-04-13 src/paper4/paper4.tex
Updated paper 4 and style

@@

    \caption{\panelref{A} Time course  of key events within a single trial. In each of the two intervals, a 0.50 \si{\second} fixation period (red) precedes the lateral translation (yellow). A 1.75 \si{\second} long delay period (shown in white) separates the two intervals. After the second translation, the participant responded whether this second translation was longer or shorter than the first. \panelref{B} and \panelref{C} Top-view illustrating key events in a far versus near body-fixed fixation trial (\panelref{B}) and a near versus far world-fixed fixation trial (\panelref{C}). The first panel shows the initial fixation to the target (red cross) followed by a translation in the second panel. The third panel shows the initial fixation before the second movement interval followed by a translation in the fourth panel.}
    \label{p4:fig1}
\end{figure}

The timing of a single trial is shown in \figref{p4:fig1}A. Every trial started with the onset of a central fixation point (i.e. aligned between the eyes), at a depth of 50 or 200\si{\centi\metre}, for 0.5 \si{\second}. Subsequently, the first 1\si{\second} motion interval commenced. During this translation, the fixation point either remained world stationary (world condition), or moved along with the participant (body condition).  Subsequently a 1.75 \si{\second} delay followed in which the sled was stationary and no  fixation light was shown. Next, a central fixation point reappeared at the other depth (50 or 200\si{\centi\metre}) than was used in the first interval. After 0.5 \si{\second}, the second 1 \si{\second} translation interval started, with the same fixation condition as in the first interval. After this second interval, the participant responded whether the second displacement was perceived longer or shorter than the first by moving a 1-dimensional joystick away from (longer) or towards (shorter) the body. Top-view illustrations of example body-centered and world-centered trials are shown in \figref{p4:fig1} panels B and C respectively.

@@

    \caption{\panelref{A} Actual (solid lines) and ideal (dashed lines) eye movement traces of one participant in the body-fixed (brown and orange) and world-fixed conditions (purple and pink). Gaze was directed at a near (brown and purple) or far (orange and pink) target. All traces shown are for 10m reference displacements. \panelref{B} Normalized eye position for each participant (\textpm 95\% confidence interval) at the end of the translation interval for the near and far world fixed targets (purple and pink respectively) as well as the near and far body fixed fixation targets (brown and orange respectively). In addition, the average \textpm SE across all participants is shown. Zero indicates the eyes remained stationary relative to the body, and one indicates the eyes followed the near world fixed target perfectly.}
    \label{p4:fig2}
\end{figure}

We first investigated the ability of participants to fixate body and world stationary targets. \figref{p4:fig2}A depicts exemplar eye traces for the 10cm reference translation with nearby and far fixation points in both the body and world condition. Changes in gaze are largely absent in the body near and body far conditions (brown and orange traces respectively), as required. During the world conditions, the eye excursions were large when fixating nearby targets and small when fixating far away ones (purple and pink traces respectively), which reflects the geometrical constraints.

We normalized the eye movement data by taking the average ratio between the measured eye excursion and the geometrically required eye displacement were the target world stationary. \figref{p4:fig2}B shows that these normalized eye displacements are about zero during body fixed fixations (brown and orange) and close to one in the world fixed fixations (purple and pink data points), for all participants. The question is whether these eye movements are inversely scaled by fixation depth in order to interpret them as linear self-motion cues.

@@

	\caption{\panelref{A} and \panelref{B} Psychometric curves (colored lines) and associated binned data (circles) for one participant. Circle size represents the amount of trials within the bin. Psychometric curves before collapsing across reference order are shown as gray lines. \panelref{A} World-fixed condition (purple) while fixation was either near (dark) or far (light). \panelref{B} Body-fixed condition (brown) while fixation was either near (dark) or far (light).
	\panelref{C} and \panelref{D} Individual and average points of subjective equality (PSEs). Color scheme matches panels A and B.
	}
	\label{p4:fig3}
\end{figure}

\figref{p4:fig3} illustrates psychophysical data on self-motion perception of a single participant for the two-fixation depths in both the world (\figref{p4:fig3}A) and body condition (\figref{p4:fig3}B). Lighter and darker colors indicate which fixation depth was the reference translation (see figure legend). The influence of fixation depth is characterized by a shift of the psychometric functions relative to the 10 \si{\centi\metre} reference translation (i.e. the PSE). For example, the rightward shift of the pink curve in \figref{p4:fig3}A, representing the world-condition, means that with a far target a longer translation (\about15 \si{\centi\metre}) was required to be perceived equivalent to a 10 \si{\centi\metre} reference translation with nearby fixation. Likewise, the leftward shift of the purple curve indicates that a shorter translation with near fixation (\about6 \si{\centi\metre}) is required to be perceived the same as the 10 \si{\centi\metre} reference translation with far fixation. Together, these opposite biases suggest that translations are perceived shorter for fixations further away. For the body condition, no shift of the psychometric curves is visible (\figref{p4:fig3}B), indicating that fixation depth (i.e. near versus far) has no effect in absence of eye movements.

Similar results were obtained for all participants, as shown by the individual PSEs (right column of \figref{p4:fig3}). Statistical significance of the effects of fixation depth was evaluated by comparing PSEs for the two fixation depths using a paired t-test. PSEs  differed significantly between the two fixation depths in the world condition, $t(7) = 5.42$, $p < 0.01$ (\figref{p4:fig3}C), but not in the body condition, $t(7) = -1.17$, $p = 0.28$ (\figref{p4:fig3}D), confirming the single subject results (\figref{p4:fig3}A, B). Thus, increasing fixation depth does not influence self-motion perception during body-stationary fixations, but causes self-motion to be perceived as shorter during world-stationary fixations.

@@

We used a simple linear model to quantify to what extent eye movements are scaled by fixation depth in order to serve as a linear self-motion cue (see \nameref{p4:sec:methods}). This model describes the perceived translation distance as a weighted average of a vestibular estimate, equal to the actual translation, and an oculomotor-based estimate. The latter estimate depends on the eye excursion which should be scaled by fixation depth to serve as a valid cue. Our model contained two weighting parameters, $\alpha_{50}$ and $\alpha_{200}$, one per fixation depth (see \tabref{p4:tab2} for best-fit values). Using these parameters we predicted the PSEs, i.e. $m_p$ in \eqnref{p4:eq5}, and plotted them against the actually observed PSEs in \figref{p4:fig4}. The positive correlation ($\rho = 0.xxxxxxxxxxxxxxxxxxxx-need-to-setup-matlab-for-this-xxxxxxxxxxxxxxxxxxx$, $p < 0.01$) between observed and predicted PSEs shows that our simple model does reasonably well in predicting perceptual performance.

@@

% Model results
To quantify the relative depth-dependent scaling of eye movements for nearby and far away fixation targets, we fitted a straightforward linear model to the perceptual responses based on the oculomotor behavior across four conditions. While two participants show partial scaling, the other six participants did not show any sign of scaling. Thus, we conclude that oculomotor signals provide a robust cue to translation perception, but are not properly scaled by fixation depth.

@@

In our previous experiment we compared translation perception with body-fixed versus world-fixed fixations at near depth (50 \si{\centi\metre}) only \cite{clemens2015a}. \figref{p4:fig6} shows how well our $\alpha_{50}$ parameter explains the data in our previous paper. The positive correlation between the actual PSEs in the previous study and those predicted using the model of the present paper ($\rho = 0.60$, $p = 0.06$) adds confidence to the parameter values presented here. The average difference between the values found here and those reported previously (see \tabref{p4:tab2}) is 12 \textpm 8 percent-points, which is relatively small given the independent measurements.

@@

It is important to point out that while the vestibular signal and thus its noise level is constant for a given translation distance, the noise in the associated oculomotor estimate might change with fixation distance because the magnitude of the eye movement is modulated by fixation depth and may show signal-dependent noise.

>>>>

\label{p4}

@@

    \caption{\panel{A} Time course  of key events within a single trial. In each of the two intervals, a 0.50 \si{\second} fixation period (red) precedes the lateral translation (yellow). A 1.75 \si{\second} long delay period (shown in white) separates the two intervals. After the second translation, the participant responded whether this second translation was longer or shorter than the first. \panel{B} and \panel{C} Top-view illustrating key events in a far versus near body-fixed fixation trial (\panel{B}) and a near versus far world-fixed fixation trial (\panel{C}). The first panel shows the initial fixation to the target (red cross) followed by a translation in the second panel. The third panel shows the initial fixation before the second movement interval followed by a translation in the fourth panel.}
    \label{p4:fig1}
\end{figure}

The timing of a single trial is shown in \panelref{p4:fig1}{A}. Every trial started with the onset of a central fixation point (i.e. aligned between the eyes), at a depth of 50 or 200\si{\centi\metre}, for 0.5 \si{\second}. Subsequently, the first 1\si{\second} motion interval commenced. During this translation, the fixation point either remained world stationary (world condition), or moved along with the participant (body condition).  Subsequently a 1.75 \si{\second} delay followed in which the sled was stationary and no  fixation light was shown. Next, a central fixation point reappeared at the other depth (50 or 200\si{\centi\metre}) than was used in the first interval. After 0.5 \si{\second}, the second 1 \si{\second} translation interval started, with the same fixation condition as in the first interval. After this second interval, the participant responded whether the second displacement was perceived longer or shorter than the first by moving a 1-dimensional joystick away from (longer) or towards (shorter) the body. Top-view illustrations of example body-centered and world-centered trials are shown in \panelref{p4:fig1}{B} and \hyperref[p4:fig1]{C} respectively.

@@

    \caption{\panel{A} Actual (solid lines) and ideal (dashed lines) eye movement traces of one participant in the body-fixed (brown and orange) and world-fixed conditions (purple and pink). Gaze was directed at a near (brown and purple) or far (orange and pink) target. All traces shown are for 10m reference displacements. \panel{B} Normalized eye position for each participant (\textpm 95\% confidence interval) at the end of the translation interval for the near and far world fixed targets (purple and pink respectively) as well as the near and far body fixed fixation targets (brown and orange respectively). In addition, the average \textpm SE across all participants is shown. Zero indicates the eyes remained stationary relative to the body, and one indicates the eyes followed the near world fixed target perfectly.}
    \label{p4:fig2}
\end{figure}

We first investigated the ability of participants to fixate body and world stationary targets. \panelref{p4:fig2}{A} depicts exemplar eye traces for the 10cm reference translation with nearby and far fixation points in both the body and world condition. Changes in gaze are largely absent in the body near and body far conditions (brown and orange traces respectively), as required. During the world conditions, the eye excursions were large when fixating nearby targets and small when fixating far away ones (purple and pink traces respectively), which reflects the geometrical constraints.

We normalized the eye movement data by taking the average ratio between the measured eye excursion and the geometrically required eye displacement were the target world stationary. \panelref{p4:fig2}{B} shows that these normalized eye displacements are about zero during body fixed fixations (brown and orange) and close to one in the world fixed fixations (purple and pink data points), for all participants. The question is whether these eye movements are inversely scaled by fixation depth in order to interpret them as linear self-motion cues.

@@

	\caption{\panel{A} and \panel{B} Psychometric curves (colored lines) and associated binned data (circles) for one participant. Circle size represents the amount of trials within the bin. Psychometric curves before collapsing across reference order are shown as gray lines. \panel{A} World-fixed condition (purple) while fixation was either near (dark) or far (light). \panel{B} Body-fixed condition (brown) while fixation was either near (dark) or far (light).
	\panel{C} and \panel{D} Individual and average points of subjective equality (PSEs). Color scheme matches panels A and B.
	}
	\label{p4:fig3}
\end{figure}

\figref{p4:fig3} illustrates psychophysical data on self-motion perception of a single participant for the two-fixation depths in both the world (\panelref{p4:fig3}{A}) and body condition (\panelref{p4:fig3}{B}). Lighter and darker colors indicate which fixation depth was the reference translation (see figure legend). The influence of fixation depth is characterized by a shift of the psychometric functions relative to the 10 \si{\centi\metre} reference translation (i.e. the PSE). For example, the rightward shift of the pink curve in \figref{p4:fig3}A, representing the world-condition, means that with a far target a longer translation (\about15 \si{\centi\metre}) was required to be perceived equivalent to a 10 \si{\centi\metre} reference translation with nearby fixation. Likewise, the leftward shift of the purple curve indicates that a shorter translation with near fixation (\about6 \si{\centi\metre}) is required to be perceived the same as the 10 \si{\centi\metre} reference translation with far fixation. Together, these opposite biases suggest that translations are perceived shorter for fixations further away. For the body condition, no shift of the psychometric curves is visible (\panelref{p4:fig3}{B}), indicating that fixation depth (i.e. near versus far) has no effect in absence of eye movements.

Similar results were obtained for all participants, as shown by the individual PSEs (right column of \figref{p4:fig3}). Statistical significance of the effects of fixation depth was evaluated by comparing PSEs for the two fixation depths using a paired t-test. PSEs  differed significantly between the two fixation depths in the world condition, $t(7) = 5.42$, $p < 0.01$ (\panelref{p4:fig3}{C}), but not in the body condition, $t(7) = -1.17$, $p = 0.28$ (\panelref{p4:fig3}{D}), confirming the single subject results (\panelref{p4:fig3}{A and B}). Thus, increasing fixation depth does not influence self-motion perception during body-stationary fixations, but causes self-motion to be perceived as shorter during world-stationary fixations.

@@

We used a simple linear model to quantify to what extent eye movements are scaled by fixation depth in order to serve as a linear self-motion cue (see \nameref{p4:sec:methods}). This model describes the perceived translation distance as a weighted average of a vestibular estimate, equal to the actual translation, and an oculomotor-based estimate. The latter estimate depends on the eye excursion which should be scaled by fixation depth to serve as a valid cue. Our model contained two weighting parameters, $\alpha_{50}$ and $\alpha_{200}$, one per fixation depth (see \tabref{p4:tab2} for best-fit values). Using these parameters we predicted the PSEs, i.e. $m_p$ in \eqnref{p4:eq5}, and plotted them against the actually observed PSEs in \figref{p4:fig4}. The positive correlation ($\rho = 0.92$, $p < 0.01$) between observed and predicted PSEs shows that our simple model does reasonably well in predicting perceptual performance.

@@

% Model results
To quantify the relative depth-dependent scaling of eye movements for nearby and far away fixation targets, we fitted a straightforward linear model to the perceptual responses based on the oculomotor behavior across four conditions. While two participants show partial scaling, the other six participants did not show any sign of scaling. Thus, we conclude that while oculomotor signals provide a robust cue to translation perception, they are not properly scaled by fixation depth.

@@

In our previous experiment we compared translation perception with body-fixed versus world-fixed fixations at near depth (50 \si{\centi\metre}) only \cite{clemens2015a}. \figref{p4:fig6} shows how well our $\alpha_{50}$ parameter explains the data in our previous paper. The positive correlation between the actual PSEs in the previous study and those predicted using the model of the present paper ($\rho = 0.60$, $p = 0.06$) adds confidence to the parameter values presented here. The average difference between the values found here and those reported previously (see \tabref{p4:tab2}) is 12 \textpm 8 percent-points, which is relatively small given the indpendent measurements.

@@

It is important to point out that while the vestibular signal and thus noise is constant for a given translation distance, the noise in the associated oculomotor estimate might change with fixation distance because the magnitude of the eye movement is modulated by fixation depth and may show signal-dependent noise.
